---
title: "Three Genes Convergence Plots"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'png')
```

This document contains plots and statistics for evaluating the convergence of 3_genes MrBayes runs performed in this paper. Assessment of convergence used the web address http://danwarren.net/plot-comparisons.html written by the authors of the package **RWTY**. 

It seems that both 3_genes runs have converged. The approximate sampling Estimated Sample Size (ESS) and topology ESS are both above 200 (standard rule of thumb). The tree topology trace shows well-mixed chains and a clear optimum, cumulative split frequencies seem to have become fairly stable, and sliding window split frequencies show large jumps, but a clear search of the tree space. Both runs also seem to have converged on similar posterior probabilities, as shown by the split frequency comparisons. The average standard deviation of split frequencies (ASDSF) is below 0.01 and shows a consistent decrease, as expected. Tree space plots show both runs searching similar areas of tree space and well-mixed chains. The topological autocorrelation plots also appear to have reached a plateau for each run.

## Analysis code

```{r code}
library(rwty)
library(ape)

# Pull in the trees (only need to indicate a folder)
my.trees <- load.multi("3_genes_longer_run", format='mb')

# Set burn in
burnin_val <- 25000

# Analysis of trees using **RWTY**
my.trees.rwty <- analyze.rwty(my.trees, burnin=burnin_val, fill.color='LnL')

# Approximate ESS for topologies
topological.approx.ess(my.trees,burnin=burnin_val)
```

\newpage
## Plots

### Estimated Sample Size

```{r ESS}
makeplot.pseudo.ess(my.trees, burnin = 2500)
```

\newpage
### Parameter plot

```{r LnL, fig.align='center', message=FALSE, warning=FALSE, fig.height = 3, fig.width = 5, dpi=300}
my.trees.rwty$LnL.trace[[1]]
my.trees.rwty$LnL.trace[[2]]
```

\newpage
### Topology trace plots

```{r top, fig.align='center', message=FALSE, warning=FALSE, fig.height = 3, fig.width = 5, dpi=300}
my.trees.rwty$topology.trace.plot[[1]]
my.trees.rwty$topology.trace.plot[[1]]
```

\newpage
### Split frequency plots 

```{r split.cumulative, fig.align='center', message=FALSE, warning=FALSE, fig.height = 3, fig.width = 5, dpi=300}
my.trees.rwty$splitfreqs.cumulative.plot
```

```{r split.sliding, fig.align='center', message=FALSE, warning=FALSE, fig.height = 3, fig.width = 5, dpi=300}
my.trees.rwty$splitfreqs.sliding.plot
```

```{r splitmatrix, fig.align='center', message=FALSE, warning=FALSE, fig.height = 3, fig.width = 5, dpi=300}
my.trees.rwty$splitfreq.matrix
my.trees.rwty$asdsf.plot
```

\newpage
### Tree space plots 

```{r treespace,fig.align='center', message=FALSE, warning=FALSE, fig.height = 3, fig.width = 5, dpi=300}
my.trees.rwty$treespace.heatmap
```

```{r treespace.points,fig.align='center', message=FALSE, warning=FALSE, fig.height = 3, fig.width = 5, dpi=300}
my.trees.rwty$treespace.points.plot
```

\newpage
### Autocorrelation plots

```{r autocorr,fig.align='center', message=FALSE, warning=FALSE, dpi=300}
my.trees.rwty$autocorr.plot
```
